Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2023-02-16 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd
just CF vs non-CF

# Chloride groups

We performed receiver operating characteristic (ROC) analysis to find a sweat sodium cutoff that distinguishes between chloride groups. Only the first sweat test from each participant was used. Because the groups are based on chloride value, Na/Cl ratio is correlated with the outcome so Na alone was used instead. Also, ROC analyses generally assume a binary outcome (i.e. looking for a cutoff to distinguish between two groups), so

In the following results, "Low" indicates a sweat chloride under 30, "Medium" indicates a sweat chloride of 30 - 59, and "High" indicates a sweat chloride greater than or equal to 60.

```{r}
# Three groups
t1_df$Group <- cut(t1_df$Cl, breaks = c(0, 30, 60, Inf), right = F, labels = c("Low", "Medium", "High"))
# Low vs. medium or high
t1_df$Low <- t1_df$Group
levels(t1_df$Low) <- c("Low", "Medium or High", "Medium or High")
t1_df$Low <- relevel(t1_df$Low, ref = "Medium or High")
# High vs. medium or low
t1_df$High <- t1_df$Group
levels(t1_df$High) <- c("Low or Medium", "Low or Medium", "High")
t1_df$High <- relevel(t1_df$High, ref = "Low or Medium")
```

## ROC analysis

```{r}
roc_low <- roc(response = t1_df$Low, predictor = t1_df$Na, quiet = T)
coords_low <- round(coords(roc_low, "best", best.method = "youden"), 3)
roc_high <- roc(response = t1_df$High, predictor = t1_df$Na, quiet = T)
coords_high <- round(coords(roc_high, "best", best.method = "youden"), 3)

plot(roc_low, print.auc = TRUE, col = "blue")
plot(roc_high, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
```

The blue curve shows the ROC for distinguishing between low vs. medium or high, and the green is the ROC curve for distinguishing between low or medium vs. high. The optimal threshold for distinguishing between low vs. medium or high was `r coords_low$threshold` (sensitivity = `r coords_low$sensitivity`, specificity = `r coords_low$specificity`) based on the Youden index. The optimal threshold for distinguishing between high vs. low or medium was `r coords_high$threshold` (sensitivity = `r coords_high$sensitivity`, specificity = `r coords_high$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_low$auc,roc_high$auc)),3)`.

## Histogram by group

Vertical red lines indicate the thresholds identified above.

```{r warning=FALSE}
ggplot(t1_df[!is.na(t1_df$Group), ], aes(Na, fill = Group)) +
  geom_histogram(alpha = 0.3, position = "identity", binwidth = 1) +
  geom_vline(xintercept = coords_low$threshold, color = "red") +
  geom_vline(xintercept = coords_high$threshold, color = "red") +
  ylab("Count") +
  theme_bw()
```

>>>>

# ROC analyses

We performed receiver operating characteristic (ROC) analysis to find a sweat sodium to sweat chloride cutoff that distinguishes between groups. Only the first sweat test from each participant was used.

Due to low numbers of people with CFTR-related disorder or CRMS, this analysis focused on CF vs. non-CF until we receive additional data from NJH.

## Diagnosis

### All participants

```{r}
# Exclude those with CFTR-related disorder or CRMS for now
roc_df = t1_df %>% filter(Diagnosis %in% c("Cystic Fibrosis","Non-CF"))
roc_df$Diagnosis = factor(roc_df$Diagnosis)
roc_df$Diagnosis = relevel(roc_df$Diagnosis,ref = "Non-CF")

roc_cf <- roc(response = roc_df$Diagnosis, predictor = roc_df$`Na/Cl Ratio`, quiet = T)
coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)

plot(roc_low, print.auc = T,print.thres=T)
```

The optimal Na/Cl threshold for distinguishing between CF and non-CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index, with and AUC of `r round(as.numeric(roc_cf$auc),3)`.

###
